\documentclass{article}

\usepackage{url}

\title{Review Reply to JMLR Machine Learning Open Source Software Submission (14-324):\\
BayesPy: Variational Bayesian Inference in Python}

\author{Jaakko Luttinen\\
  \url{jaakko.luttinen@aalto.fi}\\
    Aalto University, Finland}

\begin{document}

\maketitle

\section{Responses to the Reviewers of Submission 14-324 (v2)}

Again, I want to express my gratitude to the action editor and the reviewers for
their evaluation and comments.  I have taken into account the reviews and made a
minor revision.  In addition, I have switched to a more permissive license, that
is, from GPLv3 to MIT and released a new version (0.3.3).  Below are my comments
to the issues listed by the action editor:

\begin{itemize}

\item \underline{resolve the different factorisation}

  In the comparison to Infer.NET, I have now also used the same factorization
  for Infer.NET as for BayesPy.  This was not presented previously for a fair
  comparison because Infer.NET was performing extremely badly with the
  factorization used in BayesPy.  Perhaps, Infer.NET is optimized for the full
  factorization.  In any case, the results now summarize both the full
  factorization and the other factorization for Infer.NET.  In addition, because
  of some other updates to BayesPy, the experiments were run again, thus the
  results slightly differ from the previous results.


\item \underline{make the typographic changes}

  I thank for the language suggestions.  I have made the suggested typographic
  changes.
  

\item \underline{ensure that the install is smooth}

  I have now checked that the example in the paper works with the latest release
  in PyPI and the development version in GitHub.

  
\end{itemize}

\end{document}


%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-master: t
%%% End: 
